---
title: "Data Science with Neo4j: Recommendations"
author: "Estefania Mancini"
date: "5/25/2020"
output:
  html_document:
    code_folding: hide
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    number_sections: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

# Data Science with Neo4j
  
This course introduces you to using Neo4j as part of your Data Science and Machine Learning workflows. We’re going to learn how to do this with the help of the aminer.org citation dataset. This dataset contains papers, authors, and citations from DBLP – a computer science bibliography website. This course is intended for data scientists and data analysts. This self-paced training should take you three hours to complete if you perform all of the hands-on exercises in the course.

We have set up a discussion area in our Neo4j Community Site, if you run into problems in the course and need assistance. You should register on the community site where you can view other questions and answers for students taking our online training courses. The Neo4j Community Site is an excellent resource for answering many types of questions posed by other users of Neo4j.

There are four modules in this course. Most modules have hands-on exercises you should complete and all modules have a set of review questions at the end. The hands-on exercises in this course can be completed in a Neo4j Sandbox or in Neo4j Desktop. We have provided an estimate of how long each module should take you to complete if you perform the hands-on exercises.  

## Setting Up Your Development Environment
To configure this citation database, I followed this forum:
 < https://community.neo4j.com/t/download-datasets/11723/6 >
 
I replicate my local neo4j folder, and started a second neo4j machine to avoid messy. Then I follow this instructions:

* Here is the Citations dataset:
https://s3.amazonaws.com/neo4j-sandbox-usecase-datastores/v3_4/citations.db.zi
* Unzip the database zip file.
* Start the database.
* Stop the database.
* Copy the folder where you unzipped the database to the databases folder.
* (Delete the graph.db folder)
* Rename the to graph.db.
* Start the database.

Once it is up, you can connect using http://localhost:7474/browser/

## Import data using R:
I use this package: https://github.com/neo4j-rstats/neo4r

```{r Load, echo = TRUE}
library(neo4r)
con <- neo4j_api$new(
  url = "http://localhost:7474", 
  user = "neo4j", 
  password = ""
)
```


## Check connection and some specificities

```{r check1, echo = TRUE}
con$ping()
con$get_version()
con$get_constraints()
con$get_labels()
con$get_relationships()
con$get_index()
```

# Collaborative Filtering
Collaborative filtering is based on the idea that people like things similar to other things they like, and things that are liked by other people with similar taste.
In the Citations dataset there are a couple of different recommendations that we can generate.
uthors may be interested in reading other papers written by their coauthors. We can also recommend future collaborators by finding the people that their coauthors have collaborated with.


## Exercise 1: Coauthor Collaborative Filtering with Cypher
You will write recommendation queries to suggest potential collaborators.

### Finding popular authors
Since we're going to make collaborator suggestions find authors who have written the most articles so that we have some data to work with.
(Data is ordered by popularity)

```{r authors, echo = TRUE}
authors<-
as.data.frame(
  ' MATCH (author:Author)
    RETURN author.name, size((author)<-[:AUTHOR]-()) AS articlesPublished
    ORDER BY articlesPublished DESC
    LIMIT 10' %>%   call_neo4j(con))
head(authors)
```

### Pick one of these authors and retrieve the articles they've published and how many citations they've received:

```{r authors1, echo = TRUE}
authorName<-"Peter G. Neumann"
Peter<-as.data.frame('MATCH (:Author {name: "Peter G. Neumann"})<-[:AUTHOR]-(article)
RETURN article.title AS article, article.year AS year, size((article)<-[:CITED]-()) AS citations
ORDER BY citations DESC
LIMIT 20' %>%   call_neo4j(con))
head(Peter)
```

### Find the author's collaborators:

```{r collaborators, echo = TRUE}
collab<-as.data.frame('MATCH (:Author {name: "Peter G. Neumann"})<-[:AUTHOR]-(article)-[:AUTHOR]->(coauthor)
RETURN coauthor.name AS coauthor, count(*) AS collaborations
ORDER BY collaborations DESC
LIMIT 10' %>%   call_neo4j(con))
head(collab)
```

### How would you suggest some future collaborators for this author?

* One way is by looking at the collaborators of their collaborators!
Each of these people have collaborated with someone that Peter has worked with before, so they might be able to do an introduction.

```{r NewCollaborators, echo = TRUE}
newCollab<-as.data.frame('MATCH (author:Author {name: "Peter G. Neumann"})<-[:AUTHOR]-(article)-[:AUTHOR]->(coauthor),
(coauthor)<-[:AUTHOR]-()-[:AUTHOR]->(coc)
WHERE not((coc)<-[:AUTHOR]-()-[:AUTHOR]->(author)) AND coc <> author      
RETURN coc.name AS coauthor, count(*) AS collaborations
ORDER BY collaborations DESC
LIMIT 10'  %>%   call_neo4j(con))

head(newCollab)
```


### Exercise: Can you find the top 20 **suggested collaborators** for **Brian Fitzgerald** instead of 'Peter G. Neumann'?

```{r NewCollaboratorsEX, echo = TRUE}
newCollab<-as.data.frame('MATCH (author:Author {name: "Brian Fitzgerald"})<-[:AUTHOR]-(article)-[:AUTHOR]->(coauthor),
(coauthor)<-[:AUTHOR]-()-[:AUTHOR]->(coc)
WHERE not((coc)<-[:AUTHOR]-()-[:AUTHOR]->(author)) AND coc <> author      
RETURN coc.name AS coauthor, count(*) AS collaborations
ORDER BY collaborations DESC
LIMIT 20'  %>%   call_neo4j(con))
colnames(newCollab)<-c("person","times")
head(newCollab)
```

### How many of these potential collaborators have collaborated with Brian's collaborators more than 3 times?

```{r NewCollaboratorsEX2, echo = TRUE}
top<-as.data.frame('MATCH (author:Author {name: "Brian Fitzgerald"})<-[:AUTHOR]-(article)-[:AUTHOR]->(coauthor),
      (coauthor)<-[:AUTHOR]-()-[:AUTHOR]->(coc)
WHERE not((coc)<-[:AUTHOR]-()-[:AUTHOR]->(author)) AND coc <> author      
WITH coc, count(*) AS collaborations
WHERE collaborations > 3
RETURN count(*)' %>%   call_neo4j(con))
top
```

* Alternatively, using R:
```{r NewCollaboratorsEX3, echo = TRUE}
print(length(which(newCollab$times>3)))
```
## Anonymous graphs
The typical workflow when using the GDS library is to create a graph and store it in the catalog. This is useful to minimize reads from Neo4j and to run an algorithm with various settings or several algorithms on the same graph projection.

However, if you want to quickly run a single algorithm, it can be convenient to use an anonymous projection. The syntax is similar to the ordinary syntax for gds.graph.create, described here. It differs however in that relationship projections cannot have more than one property. Moreover, the nodeProjection and relationshipProjection arguments are named and placed in the configuration map of the algorithm


## Exercise 2: Article recommendations with Personalized PageRank
Centrality algorithms are used to determine the importance of distinct nodes in a network. The Neo4j GDS library includes the following centrality algorithms, grouped by quality tier:
* Production-quality:
  + Page Rank

* Alpha
  + ArticleRank
  + Betweenness Centrality
  + Closeness Centrality
  + Degree Centrality
  + Eigenvector Centrality

In this exercise, you will gain experience using the PageRank algorithm, understand the difference between PageRank and PPR, and use PPR to suggest relevant articles to an author.

### PageRank
PageRank is an algorithm that measures the **transitive influence** or **connectivity of nodes**. It can be computed by either iteratively distributing one node’s rank (originally based on degree) over its neighbors or by randomly traversing the graph and counting the frequency of hitting each node during these walks. The Page Rank algorithm measures the importance of each node within the graph, based on the number incoming relationships and the importance of the corresponding source nodes. The underlying assumption roughly speaking is that a page is only as important as the pages that link to it.

(Check: https://neo4j.com/docs/graph-data-science/current/appendix-a/ for sintaxis)

Run this PageRank code over the whole graph to find out the **most influential article** in terms of **citations**:

```{r query1, echo = TRUE}
pr<-as.data.frame('CALL gds.pageRank.stream("myGraph", { maxIterations: 20, dampingFactor: 0.85 })
YIELD nodeId, score
RETURN gds.util.asNode(nodeId).title AS name, score
ORDER BY score DESC, name ASC LIMIT 25'%>% call_neo4j(con))
head(pr)
```

### Personalized PageRank
Personalized PageRank is a variant of PageRank that allows us to find influential nodes based on a set of source nodes.

For example, rather than finding the overall most influential articles, we could instead, find the most influential articles with respect to a given author. Execute this code to use a personalized PageRank algorithm:

```{r query2, echo = TRUE}
pr2<-as.data.frame('MATCH (a:Author {name: "Peter G. Neumann"})
<-[:AUTHOR]-(article)-[:CITED]->(other)
WITH collect(article) + collect(other) AS sourceNodes
CALL gds.pageRank.stream("myGraph", { maxIterations: 20, dampingFactor: 0.85 })
YIELD nodeId, score 
RETURN gds.util.asNode(nodeId).title AS article, score
ORDER BY score DESC LIMIT 10' %>% call_neo4j(con))
head(pr2)
```

### Topic Sensitive Search
You can also use Personalized PageRank to do 'Topic Specific PageRank'.
When an author is searching for articles to read, they want that search to take themselves as authors into account. Two authors using the same search term would expect to see different results depending on their area of research.
Create a full text search index on the 'title' and 'abstract' properties of all nodes that have the label 'Article'. Check that the full text index has been created

